### 2020-03-11 tex/abstract.tex
first writeup draft

\begin{abstract}
This document introduces several aspects related to the confirm activity problem in 
the NuCypher KMS network. These include the current adopted solution, its relation 
with both the fees and the inflation rewards, and three potential solutions that can be 
used to provide a proof that an Ursula has served a specific number of re-encryption 
requests within a given a period. We argue that the latter is a better option to be used 
for computing the fees that Alice paid for an Ursula in the network, but not for 
computing the inflation rewards.
\end{abstract}