### 2020-03-11 tex/future-work.tex

\section{Future Work}
\label{future-work}
Most of the future work on the confirm service activity issue 
will be dedicated to construct a concrete PCD-based solution. This 
includes the following:
\begin{itemize}
\item Understand PCD and its applications in a greater depth. This 
also requires exploring NIZK in general and SNARKs in specific.

\item Define an efficient output correctness predicate that characterizes 
the correctness of the statement that PCD will attest for. This could be 
hard for the confirm activity issue as the value of the output is dynamic 
based on the workload Ursula may receive.

\item Define a way to compose the proofs efficiently and produce 
a single proof at the end. This will be tied to the formulated predicate.

\item Argue about the correctness and security of the concrete scheme.
\end{itemize}

The above will take time, but the good news that this will be also helpful 
to the work on PPSC. So it is more like developing part of the  
background needed for PPSC.