### 2020-03-23 tex/confirm-service-activity.tex
reorganize the document

\section{Confirm Service Activity}

\input{tex/confirm-service-activity/threat-network-model}
\input{tex/confirm-service-activity/prelim}
\input{tex/confirm-service-activity/solutions}
\input{tex/confirm-service-activity/analysis}
\input{tex/confirm-service-activity/future-work}