### 2020-03-25 tex/confirm-service-activity.tex
edit confirm service activity section

\section{Confirm Service Activity}
In this section, we present potential solutions for the confirm service activity 
problem. The goal is to come up with provably secure solutions that allows an 
Ursula to prove that it has served a given number of distinct re-encryption 
requests within a given period.

The rest of this section outlines the network and threat models adopted by 
the proposed solutions, provides an overview of the underlying cryptographic primitives and 
protocols, after which it discusses the proposed solutions along with an analysis of 
their security and efficiency aspects. Lastly, the section concludes with directions of 
future work on the confirm service activity issue.

\input{tex/confirm-service-activity/threat-network-model}
\input{tex/confirm-service-activity/prelim}
\input{tex/confirm-service-activity/solutions}
\input{tex/confirm-service-activity/analysis}
\input{tex/confirm-service-activity/future-work}