### 2020-06-03 confirm-activity/tex/abstract.tex

\begin{abstract}
In this document we introduce several aspects related to the confirm activity problem in 
the NuCypher network. These include the different definitions of this problem, the 
underlying threat model, the current adopted solution, and its relation 
to both service fees and inflation rewards. In particular, we consider two flavors of the confirm activity issue, namely, confirm availability 
and confirm service activity. The former confirms that Ursula is available all the time 
and replies to all requests coming from Bob, while the latter provides a proof that an 
Ursula has served a specific number of distinct re-encryption 
requests within a given a period. We present potential solutions 
for each of these flavors and discuss at 
what stage of the network operation they are likely to be used.
\end{abstract}