### 2020-08-03 confirm-activity/tex/abstract.tex
fix intro and reorganize the paper

\begin{abstract}
In this document we study the confirm activity problem in the NuCypher network. This problem is concerned with devising secure techniques that allow a service-provider, or Ursula, to prove being online and responsive during any time period. We consider two flavors of the problem based on the network operation stage, namely, confirm availability and confirm service activity. The former is about confirming that Ursula is available all the time and replies to all Bob requests, while the latter is about proving that an Ursula has served a specific number of distinct re-encryption requests within a given period. We present several potential solutions for each of these types, discuss the characteristics of these solutions with respect to their security guarantees and efficiency, and provide recommendations to when each of them should be used.
\end{abstract}