### 2020-08-04 confirm-activity/tex/discussion.tex

\section{Discussion}
\label{discussion}
As we mentioned earlier, now in the NuCypher network, we have the naive confirm activity function calling approach, with a financial punishment of slashing the stake of any Ursula who does not log her activity periodically. Our early plans will be to replace this approach with the confirm availability solution introduced in Section~\ref{confirm-availability}.

For the confirm service activity, which will be deployed at a later of the network, which solution to adopt will be based on benchmark testing and the development vision of the NuCypher network. The ZKP-based solution (the one utilizing bulletproofs) provides several advantages including non-interactivity, as opposed to the committee-attestation one that needs interaction with the committee. It also supports computational soundness as opposed to the statistical one of the commit/challenge/open solution. Nonetheless, it could be the case that the committee-attestation and the Merkle-tree based solutions are more efficient especially if we incorporate committee attestation or sidechains to support other functionalities in the network.

Another related issue to confirm activity is the service payment strategy in the network. Until now we assume that Alice pays for the service by using the Ether she locks into the policy contract. Other arrangement may emerge in the system like having Bob pay for the requests he issues. Such an arrangement may have an implication on the confirm activity issue, and studying its impact, if it becomes part of the NuCypher network protocol, will be part pf our future work.